Ages of Planets and Stars Observed by Kepler

### 2024-04-17 paper/ms.tex
ms: add text for teff/logg split

{\it Near the main-sequence (bit 9)}---Many stars observed by Kepler
are far from the main sequence: this can be seen in color--absolute
magnitude diagrams, and in diagrams of e.g.~surface gravity as a
function of temperature (Figure~\ref{fig:stellarprops}).  This is a
problem for rotation-based age-dating because {\it i)} ``stars'' far
from the main-sequence include unresolved photometric binaries, and
{\it ii}) rotation periods of single stars far from the main-sequence
have likely evolved due to effects other than magnetic braking.  The
added complications of reddening and metallicity make imposing a
simple cut in any two dimensions challenging.  After exploring various
options, we found however that the $+1$$\sigma$ relative isochronal
uncertainty reported by \citet{Berger_2020a_catalog}, $+\sigma_{\rm
B20,iso}/t_{\rm B20,iso}$, reasonably separated the relevant region of
parameter space.  Another way of saying this is that gyrochronology is
probably less reliable for stars for which the upper limit on the
relative isochronal uncertainty is small.
We flag stars with $+\sigma_{\rm B20,iso}/t_{\rm B20,iso} < 0.4$ as
meeting this criterion.

>>>>

{\it Near the main-sequence (bit 9)}---Many stars observed by Kepler
are far from the main sequence: this can be seen in color--absolute
magnitude diagrams, and in diagrams of e.g.~surface gravity as a
function of temperature (Figure~\ref{fig:stellarprops}).  This is a
challenge for rotation-based age-dating because {\it i)} point sources
far from the main-sequence include unresolved photometric binaries,
and {\it ii}) effects other than magnetic braking influence the
rotation periods of bona fide single stars far from the main-sequence.
The added complications of reddening and metallicity make defining a
simple cut in any two dimensions challenging.  We explored various
options, and found that the $+1$$\sigma$ relative isochronal
uncertainty reported by \citet{Berger_2020a_catalog}, $+\sigma_{t,{\rm
B20,iso}}/t_{\rm B20,iso}$, reasonably separated FGK stars ``near'' and
``far'' from the main-sequence.  We therefore selected stars with
$t_{\rm B20,iso}/(+\sigma_{t,{\rm B20,iso}})\approx 3$, and fitted an
an $N^{\rm th}$-order polynomial in the $\log g$-$T_{\rm eff}$ plane
to these stars between 4100 and 5800\,K.  We let the order of the fit
vary from $N$=1 to 10, and minimizing the Bayesian Information Criterion
yielded $N$=6.  The resulting locus is shown in
Figure~\ref{fig:stellarprops}.  While the details of this procedure
are entirely arbitrary, the overall purpose is to exclude stars for
which there is photometric evidence that gyrochronology may not be
reliable.  The rotation period distribution of sources excluded in
this manner includes a larger fraction of ``rapid rotators'', likely
due to a combination of astrophysical and systematic effects.
